Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-02-29 paper/paper.tex
pass on section 4.2

\algdef{SE}[DOWHILE]{Do}{doWhile}{\algorithmicdo}[1]{\algorithmicwhile\ #1}%
\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
\algnewcommand\And{\textbf{and} }%
\begin{algorithm}[t!]
  \noindent\hspace{-25ex}
  \textbf{INPUT:} (overconstrained) seed rule $sr$ \\
  \noindent\hspace{-27.5ex} 
  \textbf{OUTPUT:} Set of minimized rules $\mathcal O$\\
%\vspace{-3ex}
\captionof{algorithm}{Minimization algorithm}\label{algo1}
\begin{algorithmic}[1]
  \State $\mathcal R \gets \{\mathit{sr}\}$\Comment{initial population
    of candidate solutions contains only the input
    rule.}\label{line-init}
  \State $\mathcal O \gets \emptyset$
  \Do\label{start-do-while}
      \State $\mathcal R' \gets \emptyset$ \Comment{new iteration. reset.}\label{init-rprime}
      \ForEach {$ro \in \mathcal R $}\Comment{analyze each candidate rule.}\label{start-foreach}
           \For {$i \gets 1$ to $N$}\Comment{mutate each rule $N$ times.}\label{start-for}
               \State $r \gets copy(ro)$      
               \State $j\gets \mathit{rand}(0,\mathit{len}(r))$
               \State $r[j]\gets 0$\Comment{discard option $j$ in $r$}
               \If{$\mathit{isNew(r)} \And \mathit{isPlausible}(r)$} 
                   \State $\mathcal R' \gets \mathcal R' \cup \{r\}$\label{line-progress}
               \EndIf
           \EndFor\label{end-for}
       \EndFor\label{end-foreach}
       \State $\mathcal R \gets \mathcal R'$\Comment{update new generation}
       \State $\mathcal O \gets \mathcal O \cup \mathcal
       R$\Comment{remember all rules}
  \doWhile{$\mathcal R\not= \{\}$}\Comment{repeat until reaching a fixpoint}\label{end-do-while}
\end{algorithmic}
\end{algorithm}

@@

Algorithm~\ref{algo1} shows the pseudocode of the \tname{}
minimization algorithm. The algorithm takes as input the
overconstrained rule produced by the reverse engineering step and
produces a set $\mathcal R$ of minimized rules. Line~\ref{line-init}
initializes the set $\mathcal{R}$ of candidate solutions with the seed
rule $sr$. We encode a rule as a bitvector where each index in the
vector represents one option; the value assigned at a given position
indicates whether or not the option is present in the rule. Each
iteration of the outer loop
(lines~\ref{start-do-while}-\ref{end-do-while}) represents one
generation of the population of individuals produced by the
search. The algorithm terminates when no rule in the current
generation of individuals could be further reduced without violating
the property that a candidate solution should only capture positive
traffic. This means that execution of one iteration has not reached
the statement declared at line~\ref{line-progress}, which adds an
element to the set $\mathcal{R'}$. Consequently, that set will be
empty at the end of one iteration of the outer
loop. Line~\ref{init-rprime} initializes the new generation of
candidate rules. The inner loop
(lines~\ref{start-foreach}--\ref{end-foreach}) iterates through each
candidate rule in $\mathcal R$. Note that these rules have been found
in the previous generation. The algorithm mutates each of these rules
for $N$ times, discarding one option of the rule each time. The call
to $\mathit{isNew}$ checks if $r$ has not been previously generated
through another path whereas the call to $\mathit{isPlausible}$ checks
if the rule only captures positive traffic. Function $\mathit{copy}$
creates a copy of the bitvector of a rule, function $\mathit{len}$
returns the length of the vector, and function $\mathit{rand}$
randomly chooses one index of the vector.

>>>>

\algdef{SE}[DOWHILE]{Do}{doWhile}{\algorithmicdo}[1]{\algorithmicwhile\ #1}%
\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
\algnewcommand\And{\textbf{and} }%
\setlength{\textfloatsep}{3pt}% Remove \textfloatsep
\begin{algorithm}[t!]
  \noindent\hspace{-25ex}
  \textbf{INPUT:} (overconstrained) seed rule $sr$ \\
  \noindent\hspace{-27.5ex} 
  \textbf{OUTPUT:} Set of minimized rules $\mathcal O$\\
%\vspace{-3ex}
\captionof{algorithm}{Minimization algorithm}\label{algo1}
\begin{algorithmic}[1]
  \State $\mathcal R \gets \{\mathit{sr}\}$\Comment{initial population
    of candidate solutions contains only the input
    rule.}\label{line-init}
  \State $\mathcal O \gets \emptyset$
  \Do\label{start-do-while}
      \State $\mathcal R' \gets \emptyset$ \Comment{new iteration. reset.}\label{init-rprime}
      \ForEach {$ro \in \mathcal R $}\Comment{analyze each candidate rule.}\label{start-foreach}
           \For {$i \gets 1$ to $N$}\Comment{mutate each rule $N$ times.}\label{start-for}
               \State $r \gets copy(ro)$\label{copy}
               \State $j\gets \mathit{rand}(0,\mathit{len}(r))$\label{choose-index}
               \State $r[j]\gets 0$\Comment{discard option $j$ in $r$}\label{delete-option}
               \If{$\mathit{isNew(r)} \And\mathit{isPlausible}(r)$} 
                   \State $\mathcal R' \gets \mathcal R' \cup \{r\}$\label{line-progress}
               \EndIf
           \EndFor\label{end-for}
       \EndFor\label{end-foreach}
       \State $\mathcal R \gets \mathcal R'$\Comment{update new generation}
       \State $\mathcal O \gets \mathcal O \cup \mathcal
       R$\Comment{remember all rules}
  \doWhile{$\mathcal R\not= \{\}$}\Comment{repeat until reaching a fixpoint}\label{end-do-while}
\end{algorithmic}
\end{algorithm}

Algorithm~\ref{algo1} shows the pseudocode for \tname{}'s rule
minimization. The algorithm takes as input an overconstrained rule
produced by the reverse engineering step and produces a set $\mathcal
O$ of minimized rules. Line~\ref{line-init} initializes the current
population of individuals (\ie{}, rules) $\mathcal{R}$ with the seed
rule $sr$. We encode a rule as a bitvector where each index in the
vector represents one option; the value assigned at a given position
indicates whether or not the option is present in the rule. One
iteration of the outer loop
(lines~\ref{start-do-while}-\ref{end-do-while}) processes the current
population of individuals $\mathcal R$ and creates the next generation
$\mathcal R'$. Line~\ref{init-rprime} initializes the population of
individuals for the next generation. The inner loop
(lines~\ref{start-foreach}--\ref{end-foreach}) iterates through each
candidate rule in $\mathcal R$. The algorithm mutates each of these
rules for $N$ times, discarding one option of the rule each
time. Line~\ref{copy} assigns a copy of the bitvector encoding a rule
in $\mathcal R$ to variable $r$. Function $\mathit{len}$ returns the
length of a vector and function $\mathit{rand}$ randomly chooses one
integer value in the range defined by its parameters. So, the
expression $\mathit{rand(0, len(r))}$, appearing at
line~\ref{choose-index}, returns an index of the option in $r$ to
delete at line~\ref{delete-option}. The call to $\mathit{isNew}$
checks if $r$ has not been previously generated through another path
whereas the call to $\mathit{isPlausible}$ checks if the rule only
captures positive traffic. If $r$ satisfies both criteria, it is added
to $\mathcal R'$, which denotes the next generation of individuals in
the search.

The algorithm terminates when no rule in the current
generation of individuals could be further reduced without violating
the property that a candidate solution should only capture positive
traffic. This means that execution of one iteration has not reached
the statement declared at line~\ref{line-progress}, which adds one
element to the set $\mathcal{R'}$. Consequently, as $\mathcal{R'}$ is
initialized empty that set will remain empty at the end of an
iteration of the outer loop.

@@
